---
title: "Polyglot Wizardry from Jeroen Ooms"
author: "Kenton Russell"
date: "July 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Jeroen Ooms ([\@opencpu](https://jeroenooms.github.io/)) provides `R` users a magical polyglot world of R, JavaScript, C, and C++.  This is my attempt to both thank him and highlight some of all that he has done.  Much of my new R depends on his work.

## Ooms' Packages

[metacran](http://www.r-pkg.org/) provides a list of all [Jeroen's CRAN packages](http://www.r-pkg.org/maint/jeroen.ooms@stat.ucla.edu).  Now, I wonder if any of his packages are in the *Top Downloads*.

### jsonlite
Let's leverage the helpful meta again from [metacran](http://www.r-pkg.org/) and very quickly get some assistance from *hint-hint* [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/index.html).


```{r}
library(jsonlite)
library(formattable)
library(tibble)
library(magrittr)

fromJSON("http://cranlogs.r-pkg.org/top/last-month/9") %>%
  {as_tibble(rank=rownames(.$downloads),.$downloads)} %>%
  rownames_to_column(var = "rank") %>%
  format_table(
    formatters = list(
      area(row=which(.$package=="jsonlite")) ~ formatter("span", style="background-color:#D4F; width:100%")
    )
  )
```

`jsonlite` is an ultra-fast reliable tool to convert and create `json` in `R`.  It's fast because like much Jeroen's work, he leverages `C`/`C++` libraries.  `shiny` and `htmlwidgets` both depend on `jsonlite`.

### V8

[`V8`](https://github.com/jeroenooms/V8) gives `R` its own embedded JavaScript engine to leverage functionality in JavaScript that might not exist in `R`.  For example, the [`WebCola`](http://marvl.infotech.monash.edu/webcola) constraint-based layout engine offers valuable technology not available within R.  Let's partially recreate the [smallgroups example](http://marvl.infotech.monash.edu/webcola/examples/smallgroups.html) all in R.  You might notice that the previously mentioned `jsonlite` is essential to this workflow.

```{r}
library(V8)
library(jsonlite)
library(scales)

ctx = new_context(global="window")

ctx$source("https://cdn.rawgit.com/tgdwyer/WebCola/master/WebCola/cola.min.js")

### small grouped example
group_json <- fromJSON(
  system.file(
    "htmlwidgets/lib/WebCola/examples/graphdata/smallgrouped.json",
    package = "colaR"
  )
)

# need to get forEach polyfill
ctx$source(
  "https://cdnjs.cloudflare.com/ajax/libs/es5-shim/4.1.10/es5-shim.min.js"
)

# code to recreate small group example
js_group <- '
// console.assert does not exists
console = {}
console.assert = function(){};

var width = 960,
  height = 500

graph = {
"nodes":[
  {"name":"a","width":60,"height":40},
  {"name":"b","width":60,"height":40},
  {"name":"c","width":60,"height":40},
  {"name":"d","width":60,"height":40},
  {"name":"e","width":60,"height":40},
  {"name":"f","width":60,"height":40},
  {"name":"g","width":60,"height":40}
],
"links":[
  {"source":1,"target":2},
  {"source":2,"target":3},
  {"source":3,"target":4},
  {"source":0,"target":1},
  {"source":2,"target":0},
  {"source":3,"target":5},
  {"source":0,"target":5}
],
"groups":[
  {"leaves":[0], "groups":[1]},
  {"leaves":[1,2]},
  {"leaves":[3,4]}
  ]
}

var g_cola = new cola.Layout()
  .linkDistance(100)
  .avoidOverlaps(true)
  .handleDisconnected(false)
  .size([width, height]);

g_cola
  .nodes(graph.nodes)
  .links(graph.links)
  .groups(graph.groups)
  .start()
'

# run the small group JS code in V8
ctx$eval(js_group)
```

Now, `WebCola` has done the hard work and laid out our nodes and links, so let's get their positions.

```{r}
nodes <- ctx$get('
  graph.nodes.map(function(d){
    return {name: d.name, x: d.x, y: d.y, height: d.height, width: d.width};
  })
')

links <- ctx$get('
  graph.links.map(function(d){
    return {x1: d.source.x, y1: d.source.y, x2: d.target.x, y2: d.target.y}
  })
')
```

Some great examples of packages employing `V8` are [`geojsonio`](http://www.r-pkg.org/pkg/geojsonio), [`lawn`](https://github.com/ropensci/lawn), [`DiagrammeRsvg`](http://www.r-pkg.org/pkg/DiagrammeRsvg), [`rmapshaper`](https://github.com/ateucher/rmapshaper), and [`daff`](https://github.com/edwindj/daff).


### rjade

We got layout coordinates above. Let's use another one of Jeroen's packages [`rjade`](https://github.com/jeroenooms/rjade) that provides [`jade`](http://jade-lang.com/) (now called [pug](https://github.com/pugjs/pug)) templates through `V8`.  `rjade` will let us build a `SVG` graph with our layout.

```{r}
library(rjade)
library(htmltools)

svg <- jade_compile(
'
doctype xml
svg(version="1.1",xmlns="http://www.w3.org/2000/svg",xmlns:xlink="http://www.w3.org/1999/xlink",width="960px",height="500px")
  each l in lines
    line(style={fill:none, stroke:"lightgray"})&attributes({"x1": l.x1, "x2": l.x2, "y1": l.y1, "y2": l.y2})
  each val in rects
    g
      rect(style={fill: fillColor})&attributes({"x": val.x - val.width/2, "y": val.y - val.height/2, "height": val.height - 6, "width": val.width - 6, rx: 5, ry: 5})
      text&attributes({"x": val.x, "y": val.y, "dy": ".2em", "text-anchor":"middle"})= val.name
'
      ,pretty=T
)(rects = nodes, lines = links, fillColor = "lightgray")

HTML(svg)
```

### rsvg

If we are not in the browser though with inline `SVG` support, we very likely will want a static image format such as `png` or `jpeg`.  Of course, Jeroen has that covered also with the crazy-speedy [`rsvg`](https://github.com/jeroenooms/rsvg).  Jeroen offers [`base64`](https://github.com/jeroenooms/base64), but in this case we will use `base64enc`, since it allows `raw`.

```{r}
library(rsvg)
library(base64enc)

graph_png <- rsvg_png(charToRaw(svg))

tags$img(src=dataURI(graph_png), mime="image/png")
```

### magick

Jeroen's newest package [`magick`](https://github.com/jeroenooms/magick) is in my mind the coolest.  `magick` gives us all the power of [`ImageMagick`](http://www.imagemagick.org/script/index.php) as easy `R` functions, and is pure wizardry.  I am still shocked that it compiled first try with absolutely no problems.

```{r}
library(magick)

graph_img <- image_read(graph_png)
wizard_img <- image_read("http://www.imagemagick.org/image/wizard.png")

images <- image_annotate(
  image_append(
    c(
      image_scale(image_crop(wizard_img, "600x600+100+100"), "100"),
      image_crop(graph_img, "400x400+200+0")
    )
  ),
  "Ooms is a Wizard!",
  size = 20,
  color = "blue",
  location = "+100+200"
)

tags$img(src=dataURI(image_write(images)), mime="image/png")
```

### commonmark

I should note that this document was assembled in `rmarkdown`.  RStudio gives us lots of tools for working with `rmarkdown`, but Jeroen gives us a powerful tool [`commonmark`](https://github.com/jeroenooms/commonmark).  Let's use it to give our readers other options for output.

```{r eval = FALSE}
library(commonmark)

rmarkdown::render("Readme.Rmd", "Readme.md", output_format="md_document")

h <- markdown_html(readLines("Readme.md"))
cat(h, file="Readme_commonmark.html")
```

```{r echo=FALSE}
tags$a(href="./Readme_commonmark.html", alt="link to commonmark version", "rendered with commonmark")
```

## Conclusion and Thanks

There are of course more packages, but I'll stop here.  Jeroen Ooms truly is a wizard, and the `R` community is extraordinarily blessed to have him.  Thanks so much Jeroen.

For even more wizardry, be sure to check out [opencpu](https://www.opencpu.org/apps.html) from Jeroen, which makes R available as a web service.
